---
title: "Analysis Phase"
author: "Rogelio Reyna"
date: "2024-08-06"
output: html_document
---

Now that we have cleaned all the data in the previous phase, we can proceed with the furhter analysis. Here we are going to create additional columns and generate other interesting components.

```{r Load the cleaned data frame}
df = read_csv("cleaned_df.csv")
```

Creating the new columns: ride_length, day_of_week, month_of_week, season.

```{r Creating new columns}
df <- df %>%
  mutate(ride_length = as.numeric(difftime(ended_at, started_at, units = "mins")))

df <- df %>%
  mutate(
    day_of_week = wday(started_at, label = TRUE), # Extract day of the week
    month_of_year = month(started_at, label = TRUE), # Extract month
    season = case_when(
      month(started_at) %in% c(3, 4, 5) ~ "Spring",
      month(started_at) %in% c(6, 7, 8) ~ "Summer",
      month(started_at) %in% c(9, 10, 11) ~ "Fall",
      month(started_at) %in% c(12, 1, 2) ~ "Winter"
    )
  )

head(df[c("started_at", "ended_at","ride_length","day_of_week","month_of_year","season")])
```

Summarizing data. Counting the number of rides from the previous year for each type of user.

```{r}
df[c("member_casual")] %>% 
  
  group_by(member_casual) %>% summarize(number_of_rides = n()) %>% 
  
  filter(number_of_rides>1) %>% 
  
  arrange(-number_of_rides)
```


Now we have our data complete, now let's figure out how is different casual members from

```{r Creating charts}

plot_bar <-  ggplot(data=df, aes(x = member_casual)) +
  geom_bar(aes(fill = member_casual)) +
  labs(title = "Total Rides by Member Type", x = "Member Type", y = "Number of Rides") +
  scale_y_continuous(labels = scales::comma)
 
ggsave(filename = "TotalRidesbyMemberType.png", plot=plot_bar, width = 8, height = 6, dpi=400)
```

Creating new bar chart.

```{r}
# Plot with counts displayed inside the bars
plot_bar <- ggplot(data = df, aes(x = member_casual, fill = member_casual)) +
  geom_bar() +  # geom_bar() automatically counts the occurrences
  geom_text(position = position_stack(vjust = 0.5), aes(label = scales::comma(..count..)), stat = "count", colour = "white", fontface = "bold") +  # Add count labels
  labs(title = "Total Rides by Member Type", x = "Member Type", y = "Number of Rides") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() + # Clean theme for better readability
  theme(text = element_text(color = "white")) + #Color outside the chart itself, Number of rides, and member type. but not the values in the axis

# Save the plot
ggsave(filename = "TotalRidesbyMemberType-theme-minimal.png", plot = plot_bar, width = 8, height = 6, dpi = 400)
```

```{r}
PieDonut(data= df, aes(member_casual), explode = 2, explodeDonut = true)
```

```{r}
df %>%
  count(member_casual) %>%
  ggplot(aes(x = "", y = n, fill = member_casual)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Proportion of Rides by Member Type")
```

```{r}
ggplot(data=df, aes(x = season, y = ride_length, fill = member_casual)) +
  geom_boxplot() +
  labs(title = "Ride Length Distribution by Season and Member Type", x = "Season", y = "Ride Length")
```
